---
title: "EDA"
author: "Jiayi Shen"
date: "3/25/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
library(caret) #featurePlot
library(plotmo)
```

#Load and tidy data
```{r}
#read data
rawdata <- read.csv("kc_house_data.csv", header = TRUE)

#inspect the structure of data
str(rawdata)
```

  `id`, `date`, `zipcode`, `lat`, `long` can be removed from the dataframe.   
  The cleaned dataset to be used in this project include one response variable `price` and additional 15 variables.  
  `view`, `sqft_basement` and `yr_renovated` are continuous or integer variables in the original dataset. For the purpose of easy interpretation in later modelling, we convert these variables to be binary. Then these variables indicate whether the house has been viewed by potential buyers, whether the house has basement or not and whether the house has been renovated or not, respectively.  


```{r}
#clean the rawdata; create a tidied dataset for analysis and modelling.
#subset data: only those with view >0.
housing = 
  rawdata %>% 
  select(-id, -date, -zipcode, -lat, -long) %>% 
  filter(view > 0, bedrooms <30) %>% 
  mutate(basement = ifelse(sqft_basement == 0, 0, 1),
         renovated = ifelse(yr_renovated == 0, 0, 1)) %>% 
  filter(basement > 0) %>% 
  select(-sqft_basement, -yr_renovated, -view, - sqft_living, -sqft_lot) 

# create training data and testing data.
trrows <- createDataPartition(p=0.8, housing$price, list = FALSE)
```

#Exploratory data analysis
## Continuous variables

```{r, error=TRUE}
# vector of response
y <- housing$price
x <- model.matrix(price~.,housing)[,-1]

# matrix of continuous predictors 
housing_continous = housing %>% select(c(price, sqft_living15, grade, sqft_above, yr_built, floors, condition, bedrooms, bathrooms, sqft_lot15 ))
x_continuous <- model.matrix(price~., housing_continous)[,-1]

# set up the theme of plotting
theme1 <- trellis.par.get()   
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)

featurePlot(x_continuous, y, plot = "scatter", labels = c("","Y"),
           type = c("p"), layout = c(3, 3))
```


```{r histogram of response}
#histogram of response `price` 
housing %>% 
  ggplot(aes(x = price)) + 
    geom_histogram(aes(y =..density..),  
                    bins = 50, colour = "black", fill = "white") +
    geom_density(alpha = .1, fill = "#FF6666") +
    geom_vline(aes(xintercept = mean(price)),   
               color = "red", linetype = "dashed", size = 1)+
   xlab("Price (After BoxCox Trans.)") +theme_classic()

#QQplot
housing %>% 
  ggplot(aes(sample= price )) + 
  stat_qq() + stat_qq_line(color = "red", size = 2)+ 
  theme_minimal()+
  labs(x = "Theoretical",y = "Sample", title="QQ plot")

```


```{r correlation plot}
par(mfrow = c(1, 1))
corrplot::corrplot(cor(x))
```

